European or American? A re-examination of the transatlantic influence over the IMF

### 2025-09-11 manuscript/manuscript.qmd
Major: fix 3rd para to avoid misinterpretation

Second, contrary to much existing literature that emphasises American dominance, US influence shows a more selective pattern of impact. While American ties do not significantly affect loan volumes, participation rates, or approval decisions, they are strongly associated with more lenient conditionality—fewer performance criteria, benchmarks, and prior actions attached to lending arrangements. This finding suggests that US power manifests not through broad institutional control over access to resources, but through targeted influence over the terms and conditions of assistance.

>>>>

Second, contrary to much existing literature that emphasises American dominance, US influence shows a more selective pattern of impact. While American ties do not significantly affect loan volumes, participation rates, or approval decisions, they are strongly associated with *fewer conditions*, that is, a reduced number of performance criteria, benchmarks, and prior actions attached to lending arrangements. This finding suggests that US power manifests not through broad institutional control over access to resources, but through targeted influence over the design of conditionality. However, this result should not be interpreted as evidence of systematically “more lenient” conditionality: our additional analyses find no evidence that American ties reduce the required fiscal adjustment imposed by IMF programmes.
